---
title: "Norad_Behaviour"
author: "Charlie Wilson"
date: "March 2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

Starting Norad behaviour analysis from scratch for Yuri in March 2021
CREW script adapted from the Domenech analysis scripts

Data are loaded from a pre-prepared concatenated data file, see the silenced chunks for loading of new data or parameters. 


```{r echo = FALSE, results = 'hide', warning = FALSE, message = FALSE}

#Setup and packages

#Additional packages:
library(ggplot2) 
library(ggpubr)
library(tidyr) 
#library(plyr) 
library(dplyr) 
library(lubridate) 
library(svDialogs)
library(RcppRoll)

```


```{r echo = FALSE, results = 'hide'}

##### USEFUL FUNCTIONS ######

# summarySE Function for plot DFs


## Summarizes data.
## Gives count, mean, standard deviation, standard error of the mean, and confidence 
## interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=TRUE, conf.interval=.95) {
    library(doBy)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=TRUE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # Collapse the data
    formula <- as.formula(paste(measurevar, paste(groupvars, collapse=" + "), sep=" ~ "))
    datac <- summaryBy(formula, data=data, FUN=c(length2,mean,sd), na.rm=na.rm)

    # Rename columns
    names(datac)[ names(datac) == paste(measurevar, ".mean",    sep="") ] <- measurevar
    names(datac)[ names(datac) == paste(measurevar, ".sd",      sep="") ] <- "sd"
    names(datac)[ names(datac) == paste(measurevar, ".length2", sep="") ] <- "N"
    
    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
    
    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult
    
    return(datac)
}


```



```{r echo = FALSE, eval = FALSE, results = 'hide'}

# Setup and count and order data
# Then extract it if needed
# Skip to next chunk if already saved

#monkey <- dlgInput("Enter monkey", Sys.info()["user"])$res
# Where are the data?
# 3 folders
# Done longhand in previous versions

path <- ("/Users/charliewilson/Dropbox/Brain Lyon/NORAD/Norad_Behaviour/YURI_80_20/"); 
savepath <- ("/Users/charliewilson/Dropbox/Brain Lyon/NORAD/Norad_Behaviour/"); 
Monkey <- c("Yuri")
Dat <- data.frame(); #Initialise big data frame
All_dates <- data.frame(); #Initialise big data frame

for(m in Monkey){

    #List and organise the data
    Session.list <- dir(path, pattern = "*.txt"); #List the files in the date
    Session.list <- dir(path); #List the files in the date
    Dates <- data.frame(Session.list); #Extract dates from filenames
    colnames(Dates) <- c("folder"); #rename File ID column
    Dates <- Dates %>% separate(folder, c("Monkey","_","Task","_2","Time","_3","Date"),sep=cumsum(c(4,1,11,1,5,1,11))); #Split the filename into separate sections
    Dates$Date <- dmy(Dates$Date); #Put Date column into date format
    Dates$fold <- Session.list; 
    Dates <- arrange(Dates, Date); #Put Dates into date order
    
    Nsession <- length(Dates$Date); # How many sessions?
    Data <- data.frame(); #Initialise big data frame

    if(Nsession >0){ #Sanity check
  
          for(isession in 1:Nsession){ #Loop on each session  
            
            fid <- Dates$fold[isession]; #Extract filename from Dates
            print(fid) #Show filename
            event.raw <- read.csv(file=paste(path,fid,"/BaseReport.csv", sep= ""),sep= "", dec =",",  header = FALSE, skip =3, fill=TRUE); #Load data from .txt file using space as separator
            colnames(event.raw) <- c("time","event"); #Name columns
            event.raw$session <- isession; #Record session number
            
            # Now some basic analysis
            event.raw$trial <- cumsum(event.raw$event == 100); #Trial number
            event.raw$block <- cumsum(event.raw$event == 8); #Block number
            
            # Which stimulus was optimal in the block
            event.raw <- event.raw %>%
              group_by(trial) %>%   #Apply this trial by trial (rather than line by line)
              dplyr:: mutate(Stim = nth(event,2)) %>%  #In every trial take teh nth entry in the Event column
              ungroup();  #Put it back with the new column
            
            # Which stim is rewarded 
            event.raw <- event.raw %>%
              group_by(trial) %>%
              dplyr:: mutate(L = nth(event,3), M = nth(event,4), R = nth(event,5)) %>%
              ungroup();
            
            # This is the end of intiial analysis of a single day
            
            #Put this session (event.raw) into Data which will contain all of the sessions
            Data <- rbind(Data,event.raw);
            
            
           rm(event.raw)
              
          }
    }
    
    #And save this section
    save(file = paste(savepath,Monkey,"_",Dates$Task[1],"_","Behav.Rdata",sep = ""),Data, Dates) # Save monkey specific data frame

    # Concatenation across monkeys and stages
    Data <- Data %>%
      mutate(monkey = Monkey)
    Dat <- rbind (Dat, Data)

    Dates <- Dates %>%
      mutate(monkey = Monkey)
    All_dates <- rbind (All_dates, Dates)
    
    rm ('Data', 'Dates', 'Session.list','isession')
  
}
    
#Save
save(file = paste(savepath,"/All_Data_PST_NORAD.Rdata",sep = ""),Dat, All_dates) # Save monkey specific data frame


``` 

First load the major datafile

```{r echo = FALSE, results = 'hide'}

# Where are the data?
savepath <- ("/Users/charliewilson/Dropbox/Brain Lyon/NORAD/Norad_Behaviour/"); 

load(file = paste(savepath,"All_Data_PST_NORAD.Rdata",sep = "")) 

```




First, let's track basic behaviour across sessions?
 
  
  
  
```{r echo = FALSE, warning = FALSE, message = FALSE}


#Count trial types and outcomes
# For multiple monkeys and/or phase versions, need to reinstate group_by(monkey, Phase, session) %>%
# Note that this removes completely any sessions where there is a zero count of one of the items

Sess.data <- Dat %>% group_by(monkey, session,event) %>% tally()
Sess.data2 <- subset(Sess.data,(event %in% c(100, 65)))
Sess.data2 <- Sess.data2 %>% pivot_wider(names_from = event, values_from = n) #Longer to do maths
Sess.data2$Proportion <- Sess.data2$`65` / Sess.data2$`100` 
#Sess.data <- Sess.data %>% pivot_wider(names_from = event, values_from = n) #Longer might be easier for some stuff

ggplot(Sess.data2,aes(x=session, y=Proportion) ) + #What data to plot
    geom_smooth(method = 'loess', span = 0.3, color = "green", size = 3) + #Add a specific plot and colour it
    geom_point(color = "blue") + #
    ggtitle("Proporton correct trials over sessions") +
    theme_bw()
    #facet_grid(monkey)

Sess.data$Code <- as.factor(Sess.data$event)
ggplot(subset(Sess.data,(event %in% c(100, 65, 66))),aes(x=session, y=n) ) + #What data to plot
    geom_line(aes(color = Code)) + #Add a specific plot and colour it
    ggtitle("Trials, rewards and nonrewards over sessions") +
    theme_bw()
    #facet_grid(monkey)

ggplot(subset(Sess.data,(event %in% c(121, 122, 123))),aes(x=session, y=n) ) + #What data to plot
    geom_line(aes(color = Code)) + #Add a specific plot and colour it
    ggtitle("Good touches to the 3 stimuli over sessions") +
    theme_bw()
    #facet_grid(monkey)    

ggplot(subset(Sess.data,(event %in% c(111, 112, 113))),aes(x=session, y=n) ) + #What data to plot
    geom_line(aes(color = Code)) + #Add a specific plot and colour it
    ggtitle("Pre-emptive touches to the 3 stimuli over sessions") +
    theme_bw()
    #facet_grid(monkey)    




```




This script now builds a trial by trial matrix to help the trial analysis, and populates it with the necessary information on each trial and block
This allows us to summarise what the monkey(s) are doing within and across blocks.




```{r echo = FALSE, warning = FALSE, message = FALSE}

# First some summary statistics from the whole trial
Dat <- Dat %>%
  dplyr::group_by(monkey,session,trial) %>%
  dplyr::mutate(Lt = sum(event == 121), Mt = sum(event == 122),Rt = sum(event == 123), Nt = sum(event == 252), NR = sum(event == 66), Rew = sum(event == 65)) %>% #Touches after lever validated & reward status
  dplyr::ungroup() 

# Extract response in the context of the currently optimal choice
Dat <- Dat %>%
  dplyr::group_by(monkey,session,trial) %>%
  dplyr::mutate(COR =
                  case_when(
                    #Choice is optimal for block
                    Lt==1 && Stim==51 ~ 1, 
                    Mt==1 && Stim==52 ~ 1, 
                    Rt==1 && Stim==53 ~ 1, 
                    TRUE ~ 0),
                Ch =
                  case_when(
                    #Single value of which stim chosen
                    Lt==1 ~ 1, 
                    Mt==1 ~ 2, 
                    Rt==1 ~ 3, 
                    TRUE ~ 0),
                Lr = case_when(Lt==1 && Rew==1 ~ 1, TRUE ~ 0),
                Mr = case_when(Mt==1 && Rew==1 ~ 1, TRUE ~ 0),
                Rr = case_when(Rt==1 && Rew==1 ~ 1, TRUE ~ 0),
                ) %>%
  dplyr::ungroup()

# Extract trap rewards, when rewarded but not on the currently optimal choice
Dat <- Dat %>%
  dplyr::group_by(monkey,session,trial) %>%
  dplyr::mutate(TRAP =
                  case_when(
                    #Rewarded but not COR
                    Rew==1 && COR==0 ~ 1, 
                    TRUE ~ 0)) %>%
  dplyr::ungroup()

# Now extract the first line of each trial
#rm ('Trials')
Trials <- Dat %>%
  dplyr::group_by(monkey,session,trial) %>%
  slice(1)
Trials <- subset(Trials, event==100) # Remove trial 0 from each 

# Before we remove the trials with no response, let's plot trial outcome overall
ggplot(Trials, aes(x=Ch)) + geom_histogram(binwidth = 1, stat="count") +
  ggtitle("Overall distribution of choices") +
  xlab("Choice made (0=no choice)") +
  theme_bw()

# Remove trials with no response...
# Allows us to track active choices
Trials <- subset(Trials, Nt!=1) 
#The trial counter is now broken, add another one
Trials <- Trials %>%
  dplyr::group_by(monkey,session) %>%
  dplyr::mutate(cTrial = row_number()) %>%
  dplyr::ungroup()

# Extract the previous trial info for Choice and Reward
Trials <- Trials %>%
  dplyr::group_by(monkey,session) %>%
  dplyr::mutate(PrevCh=lag(Ch), PrevRew=lag(Rew) ) %>%
  dplyr::ungroup() 

# Shifts in respose
Trials <- Trials %>%
  dplyr::group_by(monkey,session) %>%
  dplyr::mutate(Shift =
                  case_when(
                    #Choice is optimal for block
                    Ch==PrevCh ~ 0, 
                    TRUE ~ 1)) %>%
  dplyr::ungroup()

# How many rewards in the last 5 trials?
Trials <- Trials %>%
  dplyr::group_by(monkey, session) %>%
  dplyr::mutate(roll_sum = roll_sum(Rew, 5, align = "right", fill = NA)) %>%
  dplyr::ungroup() 

#BLOCK ANALYSES

# Within block trial & reward counter
Trials <- Trials %>%
  dplyr::group_by(monkey, session, block) %>%
  dplyr::mutate(bTrial = row_number(),bRew = cumsum(Rew)) %>%
  dplyr::ungroup() 

# %C and % R in block
S1 <- summarySE(Trials, measurevar="COR", groupvars=c("bTrial"))
S1$var <-"COR"
names(S1)[names(S1) == 'COR'] <- 'Val'
S2 <- summarySE(Trials, measurevar="Rew", groupvars=c("bTrial"))
S2$var <-"Rew"
names(S2)[names(S2) == 'Rew'] <- 'Val'
S3 <- summarySE(Trials, measurevar="Shift", groupvars=c("bTrial"))
S3$var <-"Shift"
names(S3)[names(S3) == 'Shift'] <- 'Val'
S4 <- summarySE(Trials, measurevar="TRAP", groupvars=c("bTrial"))
S4$var <-"TRAP"
names(S4)[names(S4) == 'TRAP'] <- 'Val'
S <- rbind(S1, S2, S3, S4)
rm("S1", "S2", "S3", "S4")

ggplot(S, aes(x=bTrial,y=Val,colour=var)) + #What data to plot
    geom_line(size=3,alpha = 0.5)+ 
    geom_errorbar(aes(ymin=Val-se, ymax=Val+se), width=.1) +
    ggtitle("Rewards, optimal choices, and shifts within block") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("Proportion") +
    xlab("Trials in block") +
    theme_bw()

# Across blocks within session
SS1 <- summarySE(Trials, measurevar="COR", groupvars=c("block"))
SS1$var <-"COR"
names(SS1)[names(SS1) == 'COR'] <- 'Val'
SS2 <- summarySE(Trials, measurevar="Rew", groupvars=c("block"))
SS2$var <-"Rew"
names(SS2)[names(SS2) == 'Rew'] <- 'Val'
SS3 <- summarySE(Trials, measurevar="Shift", groupvars=c("block"))
SS3$var <-"Shift"
names(SS3)[names(SS3) == 'Shift'] <- 'Val'
SS4 <- summarySE(Trials, measurevar="TRAP", groupvars=c("block"))
SS4$var <-"TRAP"
names(SS4)[names(SS4) == 'TRAP'] <- 'Val'
SS <- rbind(SS1, SS2, SS3, SS4)
rm("SS1", "SS2", "SS3", "SS4")


ggplot(SS, aes(x=block,y=Val,colour=var)) + #What data to plot
    geom_line(size=3,alpha = 0.5)+ 
    geom_errorbar(aes(ymin=Val-se, ymax=Val+se), width=.1) +
    theme_bw() +
    ggtitle("Rewards and optimal choices across blocks / within session") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("Proportion") +
    xlab("Blocks")  

# Touch distribution in block
# Note that the Nts should be zero as we have removed them for the proportion analysis
T1 <- summarySE(Trials, measurevar="Lt", groupvars=c("bTrial"))
T1$var <-"Lt"
names(T1)[names(T1) == 'Lt'] <- 'Val'
T2 <- summarySE(Trials, measurevar="Mt", groupvars=c("bTrial"))
T2$var <-"Mt"
names(T2)[names(T2) == 'Mt'] <- 'Val'
T3 <- summarySE(Trials, measurevar="Rt", groupvars=c("bTrial"))
T3$var <-"Rt"
names(T3)[names(T3) == 'Rt'] <- 'Val'
T <- rbind(T1, T2, T3)
rm("T1", "T2", "T3")

ggplot(T, aes(x=bTrial,y=Val,colour=var)) + #What data to plot
    geom_line(size=3,alpha = 0.5)+ 
    geom_errorbar(aes(ymin=Val-se, ymax=Val+se), width=.1) +
    geom_hline(aes(yintercept=0.333), colour="#990000", linetype="dashed") +
    ggtitle("Touch proportion within block") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("Proportion") +
    xlab("Trials in block") +
    theme_bw() +
    ylim(0,1)

# Touch distribution across blocks
TT1 <- summarySE(Trials, measurevar="Lt", groupvars=c("block"))
TT1$var <-"Lt"
names(TT1)[names(TT1) == 'Lt'] <- 'Val'
TT2 <- summarySE(Trials, measurevar="Mt", groupvars=c("block"))
TT2$var <-"Mt"
names(TT2)[names(TT2) == 'Mt'] <- 'Val'
TT3 <- summarySE(Trials, measurevar="Rt", groupvars=c("block"))
TT3$var <-"Rt"
names(TT3)[names(TT3) == 'Rt'] <- 'Val'
TT <- rbind(TT1, TT2, TT3)
rm("TT1", "TT2", "TT3")

ggplot(TT, aes(x=block,y=Val,colour=var)) + #What data to plot
    geom_line(size=3,alpha = 0.5)+ 
    geom_errorbar(aes(ymin=Val-se, ymax=Val+se), width=.1) +
    geom_hline(aes(yintercept=0.333), colour="#990000", linetype="dashed") +
    ggtitle("Touch proportion across blocks") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("Proportion") +
    xlab("Blocks") +
    theme_bw() +
    ylim(0,1)

# Touch distribution in block
# Note that the Nts should be zero as we have removed them for the proportion analysis
Tr1 <- summarySE(Trials, measurevar="Lr", groupvars=c("bTrial"))
Tr1$var <-"Lr"
names(Tr1)[names(Tr1) == 'Lr'] <- 'Val'
Tr2 <- summarySE(Trials, measurevar="Mr", groupvars=c("bTrial"))
Tr2$var <-"Mr"
names(Tr2)[names(Tr2) == 'Mr'] <- 'Val'
Tr3 <- summarySE(Trials, measurevar="Rr", groupvars=c("bTrial"))
Tr3$var <-"Rt"
names(Tr3)[names(Tr3) == 'Rr'] <- 'Val'
T <- rbind(Tr1, Tr2, Tr3)
rm("Tr1", "Tr2", "Tr3")

ggplot(T, aes(x=bTrial,y=Val,colour=var)) + #What data to plot
    geom_line(size=3,alpha = 0.5)+ 
    geom_errorbar(aes(ymin=Val-se, ymax=Val+se), width=.1) +
    geom_hline(aes(yintercept=0.333), colour="#990000", linetype="dashed") +
    ggtitle("Rewarded touch proportion within block") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("Proportion") +
    xlab("Trials in block") +
    theme_bw() +
    ylim(0,0.5)

```



Under what circumstances do monkeys shift choice?




```{r echo = FALSE}

# Reward status when shifting

Sh <- summarySE(Trials, measurevar="roll_sum", groupvars=c("Shift"))
Sh$ShiftF <- as.factor(Sh$Shift)
levels(Sh$ShiftF) <- c("Stay", "Shift")

    ggplot(Sh, aes(x=ShiftF,y=roll_sum, group=1)) + 
    geom_errorbar(width=.1, aes(ymin=roll_sum-ci, ymax=roll_sum+ci)) +
    geom_point(shape=21, size=3, fill="red") +
    ylim(0,5) +
    ggtitle("How rewarded has the monkey been when he shifts?") +
    ylab("Rewards in previous 5 trials") +
    xlab("Shift (1) or Stay") +
    theme_bw() 
    #ggplot(subset(Trials,!(is.na(Trials$roll_sum))), aes(x=ShiftF, y=roll_sum, fill=ShiftF)) + geom_boxplot() +

# Previous trial when shifting

    Sh1 <- summarySE(Trials, measurevar="Shift", groupvars=c("PrevRew"))
    ggplot(subset(Sh1,PrevRew>=0), aes(x=as.factor(PrevRew),y=Shift, group=1)) + 
    geom_errorbar(width=.1, aes(ymin=Shift-ci, ymax=Shift+ci)) +
    geom_point(shape=21, size=3, fill="green") +
    ylim(0,0.5) +
    ggtitle("Probability of shifting given previous reward") +
    ylab("Probability of shifting") +
    xlab("Previous trial rewarded?") +
    theme_bw() 
    
```

In sum (300322)

- Yuri is rewarded now in about 60% of trials
- Consistent bais towards the left stimulus in good and bad touches
- Expected within-block learning
- Not much TOT in performance
- Shifts when has been receiving fewer rewards
- Only really shfits after an incorrect
 