---
title: "Norad_Behaviour"
author: "Charlie Wilson"
date: "March 2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

Starting Norad behaviour analysis from scratchfor Yuri in March 2021
Script adapted from the Domenech analysis scripts

Data are loaded from pre-prepared data files, see the silenced chunks for loading of new data or parameters. 


```{r echo = FALSE, results = 'hide', warning = FALSE, message = FALSE}

#Setup and packages

#Additional packages:
library(ggplot2) 
library(ggpubr)
library(tidyr) 
#library(plyr) 
library(dplyr) 
library(lubridate) 
library(svDialogs)
library(RcppRoll)

```


```{r echo = FALSE, results = 'hide'}

##### USEFUL FUNCTIONS ######

# summarySE Function for plot DFs


## Summarizes data.
## Gives count, mean, standard deviation, standard error of the mean, and confidence 
## interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=TRUE, conf.interval=.95) {
    library(doBy)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=TRUE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # Collapse the data
    formula <- as.formula(paste(measurevar, paste(groupvars, collapse=" + "), sep=" ~ "))
    datac <- summaryBy(formula, data=data, FUN=c(length2,mean,sd), na.rm=na.rm)

    # Rename columns
    names(datac)[ names(datac) == paste(measurevar, ".mean",    sep="") ] <- measurevar
    names(datac)[ names(datac) == paste(measurevar, ".sd",      sep="") ] <- "sd"
    names(datac)[ names(datac) == paste(measurevar, ".length2", sep="") ] <- "N"
    
    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
    
    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult
    
    return(datac)
}


```



```{r echo = FALSE, eval = FALSE, results = 'hide'}

# Setup and count and order data
# Then extract it if needed
# Skip to next chunk if already saved

#monkey <- dlgInput("Enter monkey", Sys.info()["user"])$res
# Where are the data?
# 3 folders
# Done longhand in previous versions

paths <- ("/Users/charliewilson/Dropbox/Brain Lyon/NORAD/ANALYSES/SESSIONS/YURI/80_20C/"); 

Monkey <- c("Yuri")
Dat <- data.frame(); #Initialise big data frame
All_dates <- data.frame(); #Initialise big data frame

for(m in Monkey){

    #List and organise the data
    Session.list <- dir(paths, pattern = "*.txt"); #List the files in the date
    Dates <- data.frame(Session.list); #Extract dates from filenames
    colnames(Dates) <- c("fid"); #rename File ID column
    Dates <- Dates %>% separate(fid, c("Monkey","_","Task","_2","Time","Date","Ext"),sep=cumsum(c(4,1,11,1,6,11,4))); #Split the filename into separate sections
    Dates$Date <- dmy(Dates$Date); #Put Date column into date format
    Dates$fid <- Session.list; 
    Dates <- arrange(Dates, Date); #Put Dates into date order
    
    Nsession <- length(Dates$Date); # How many sessions?

    #CREW TO HERE 030322
    
    Data <- data.frame(); #Initialise big data frame
    
    
    if(Nsession >0){ #Sanity check
  
          for(isession in 1:Nsession){ #Loop on each session  
            
            fid <- Dates$fid[isession]; #Extract filename from Dates
            print(fid) #Show filename
            event.raw <- read.csv(file=paste(paths,p,"/",m,"/",fid, sep= ""),sep= "", dec =",",  header = FALSE, fill=TRUE, colClasses=c(NA, NA,"NULL")); #Load data from .txt file using space as separator
            colnames(event.raw) <- c("time","event"); #Name columns
            event.raw$session <- isession; #Record session number
            
            # Now some basic analysis
            event.raw$trial <- cumsum(event.raw$event == 100); #Trial number
            
            # Which stimulus was used in the trial
            event.raw <- event.raw %>%
              group_by(trial) %>%   #Apply this trial by trial (rather than line by line)
              dplyr:: mutate(Stim = nth(event,4)) %>%  #In every trial take teh nth entry in the Event column
              ungroup();  #Put it back with the new column
            
            # Which TS
            event.raw <- event.raw %>%
              group_by(trial) %>%
              dplyr:: mutate(TS = nth(event,2)) %>%
              ungroup();
            
            # This is the end of intiial analysis of a single day
            
            #Put this session (event.raw) into Data which will contain all of the sessions
            Data <- rbind(Data,event.raw);
            
            
           rm(event.raw)
              
          }
    }
    
    #And save this section
    save(file = paste(paths,p,"/Data_Domenech_",p,"_",m,".Rdata",sep = ""),Data, Dates) # Save monkey specific data frame

    # Concatenation across monkeys and stages
    Data <- Data %>%
      mutate(monkey = m, Phase = p)
    Dat <- rbind (Dat, Data)
    
    Dates <- Dates %>%
      mutate(monkey = m, Phase = p)
    All_dates <- rbind (All_dates, Dates)
    
    rm ('Data', 'Dates', 'Session.list','isession')
  
}
    
#Save
save(file = paste(paths,"/All_Data_ST123.Rdata",sep = ""),Dat, All_dates) # Save monkey specific data frame


``` 

First load the major datafile

```{r echo = FALSE, results = 'hide'}

# Where are the data?
paths <- ("/Users/charliewilson/Dropbox/Brain Lyon/Reversible Cognition/Domenech_Analysis/Behaviour/"); 

load(file = paste(paths,"/All_Data_ST123.Rdata",sep = "")) 

```

What is in each session?
 
  
```{r echo = FALSE}


#Count trial types and outcomes
Sess.data <- Dat %>%
        group_by(monkey, Phase, session) %>%   #Apply this session by session   #was group_by(session, monkey)
        filter(event==152) %>%   #Noisy good answer
        count(event)  #Count the instances
Sess.data2 <- Dat %>%
        group_by(monkey, Phase, session) %>%   #Apply this session by session
        filter(event==153) %>%   #Noisy bad answer
        count(event)  #Count the instances
Sess.data3 <- Dat %>%
        group_by(monkey, Phase, session) %>%   #Apply this session by session
        filter(event==65) %>%   #Rewards
        count(event)  #Count the instances

# Combine the DFs and make sure it is done on the basis of "session" and "monkey" and "phase"
# Note that this removes completely any sessions where there is a zero count of one of the items
# Current case session 91 for P is removed...
Sess.dat <- merge(Sess.data, Sess.data2, by=c("session","monkey","Phase")) %>%
              merge(Sess.data3, by=c("session","monkey","Phase")) 

colnames(Sess.dat) <- c("session","monkey","Phase","xx","Noiseg","yy","Noiseb","zz","Rew"); #Rename columns
Sess.dat <- subset(Sess.dat, select= -c(xx,yy,zz)); #remove superfluous columns
remove(Sess.data, Sess.data2,Sess.data3); #remove superfluous DFs

Sess.dat <- pivot_longer(Sess.dat, -c(session, monkey, Phase), values_to = "Count", names_to = "Outcome") #Make long for plotting

ggplot(Sess.dat, aes(x=session, y=Count, color=Outcome)) + #What data to plot
    geom_point(aes(shape = monkey, color = Outcome))  + #Add a specific plot and colour it
    facet_grid(monkey ~ Phase)


Sess.dat <- Sess.dat[ which(Sess.dat$Outcome!='Rew'), ]

ggplot(Sess.dat, aes(x=session, y=Count, color=Outcome)) + #What data to plot
    geom_point(aes(shape = monkey, color = Outcome, alpha=0.3))  + #Add a specific plot and colour it
    geom_smooth() +
    facet_grid(monkey ~ Phase)
    

```

Let's build a trial by trial matrix to help the trial analysis, and populate it with the necessary information on each trial and TS

```{r echo = FALSE}

# First some summary statistics from the whole trial
Dat <- Dat %>%
  dplyr::group_by(monkey,Phase,session,trial) %>%
  dplyr::mutate(GNorm = sum(event == 150), BNorm = sum(event == 151),GTrap = sum(event == 152), BTrap = sum(event == 153), Rew = sum(event == 65)) %>%
  dplyr::ungroup() 


# And use that to extract if the response was correct or not
Dat <- Dat %>%
  dplyr::group_by(monkey,Phase,session,trial) %>%
  dplyr::mutate(COR = max(GNorm | GTrap), INC = max(BNorm | BTrap)) %>%
  dplyr::ungroup() 

# Extract response
Dat <- Dat %>%
  dplyr::group_by(monkey,Phase,session,trial) %>%
  dplyr::mutate(Resp = 
                  case_when(
                    sum(event == 61)>0 ~ 1,
                    sum(event == 62)>0 ~ 2,
                    sum(event == 63)>0 ~ 3)) %>%
  dplyr::ungroup() 

# Now extract the first line of each trial
#rm ('Trials')
Trials <- Dat %>%
  dplyr::group_by(monkey,Phase,session,trial) %>%
  slice(1)
Trials <- subset(Trials, COR>0 | INC>0) # Just trials with an outcome

# Extract the previous trial info for Stim and Rew
Trials <- Trials %>%
  dplyr::group_by(monkey,Phase,session) %>%
  dplyr::mutate(PrevSt=lag(Stim), PrevOut=lag(COR) ) %>%
  dplyr::ungroup() 

# How many rewards in the last 3 trials?
Trials <- Trials %>%
  dplyr::group_by(monkey,Phase,session) %>%
  dplyr::mutate(roll_sum = roll_sum(Rew, 4, align = "right", fill = NA)) %>%
  dplyr::ungroup() 
   
# We want a new DF with the outcomes around each trap
TrapI <- which(Trials$GTrap == 1 | Trials$BTrap == 1)

# Count trials since and to the Gtraps
# First track cycles of GTraps
Trials <- Trials %>%
  dplyr:: group_by(monkey,Phase,session) %>% 
  dplyr:: mutate(GTrapCyc = cumsum(GTrap)) %>%
  dplyr:: ungroup() 

# Then count forward and backward
Trials <- Trials %>%
  dplyr::group_by(monkey,Phase,session,GTrapCyc) %>% 
  dplyr::mutate(GTrapC = row_number()-1,GTrapCback = row_number() - (n()+1))

# Subset down to trials just around Traps and set t around Traps
#rm ('Traps', 'TrapSE')
Traps <- subset(Trials, GTrapC <=2 | GTrapCback == -1)
# The trial BEFORE each trap belongs to the GTrapCyc of the trap that comes just after for this analysis...
Traps$GTrapCyc <- lead(Traps$GTrapCyc) #Shift up 1
Traps$t <-Traps$GTrapC
Traps$t[Traps$t>2] <- -1

# Is the response the same as the Trap?
Traps <- Traps %>%
  group_by(monkey,Phase,session,GTrapCyc) %>% 
  dplyr::mutate(TrapR = nth(Resp,2), SameRT = as.integer(TrapR==Resp)) 

# Is the stimulus the same as the Trap?
Traps <- Traps %>%
  group_by(monkey,Phase,session,GTrapCyc) %>% 
  dplyr::mutate(TrapS = nth(Stim,2), SameST = as.integer(TrapS==Stim))

TrapSSame <- subset(Traps, SameST == 1)
TrapSDif <- subset(Traps, SameST == 0)
```

Use this to produce a monkey version of the figures from PD's AI manuscript

```{r echo = FALSE, warning = FALSE}

#Test <- table(TrapSSame$SameRT)

# summarySE provides the standard deviation, standard error of the mean, and a (default 95%) confidence interval
# We need the summary for the different t values -1 to 2
# But for t=2 only we need to split on the basis of PrevOut 
# The following is an ugly way of achieving this...
SS1 <- summarySE(TrapSSame, measurevar="SameRT", groupvars=c("t","PrevOut","monkey","Phase"))
SS1 <- subset(SS1, t==2 & !is.na(PrevOut),select=-PrevOut)
SS2 <- summarySE(TrapSSame, measurevar="SameRT", groupvars=c("t","monkey","Phase"))
SS2 <- subset(SS2, t<2 & t!=0)
SumSame <- rbind(SS2,SS1)

# Also for the Different stimulus...
SD1 <- summarySE(TrapSDif, measurevar="SameRT", groupvars=c("t","PrevOut","monkey","Phase"))
SD1 <- subset(SD1, t==2 & !is.na(PrevOut),select=-PrevOut)
SD2 <- summarySE(TrapSDif, measurevar="SameRT", groupvars=c("t","monkey","Phase"))
SD2 <- subset(SD2, t<2)
SumDif <- rbind(SD2,SD1)

# Reproduce figs 2a,b from the Domenech Manuscript
xxx <- c("t1-","t1-","t1+","t1+")

ggplot(SumSame, aes(x=t,y=SameRT)) + #What data to plot
    geom_point(aes(colour=monkey), size=3,alpha = 0.5)+
    geom_text(aes(label=ifelse(t>1,xxx,'')),hjust=2,vjust=-1) +
    geom_errorbar(aes(ymin=SameRT-se, ymax=SameRT+se), width=.1) +
    ylim(0,1) +
    geom_hline(aes(yintercept=0.333), colour="#990000", linetype="dashed") +
    geom_rect(aes(xmin=-0.1, xmax=0.1, ymin=0, ymax=1), color="grey", alpha=0.1) +
    ggtitle("Same Stimulus as 'Negative' Trap on t0") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("Proportion repetition same respons as 'Neg' Trap") +
    xlab("Trials since 'Neg' Trap") +
    facet_grid(monkey ~ Phase) + 
    theme_bw()
   
ggplot(SumDif, aes(x=t,y=SameRT)) + #What data to plot
    geom_point(aes(colour=monkey), size=3,alpha = 0.5)+
    geom_text(aes(label=ifelse(t>1,xxx,'')),hjust=2,vjust=-1) +
    geom_errorbar(aes(ymin=SameRT-se, ymax=SameRT+se), width=.1) +
    ylim(0,1) +
    geom_hline(aes(yintercept=0.333), colour="#990000", linetype="dashed") +
    geom_rect(aes(xmin=-0.1, xmax=0.1, ymin=0, ymax=1), color="grey", alpha=0.1) +
    ggtitle("Different Stimulus from 'Negative' Trap on t0") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("Proportion repetition same respons as 'Neg' Trap") +
    xlab("Trials since 'Neg' Trap") +
    facet_grid(monkey ~ Phase)+ 
    theme_bw()
  

# Reproduce figs 2c,d from the Domenech Manuscript

# Summarise rolling sum of same resp
Sumt1same <- summarySE(subset(TrapSSame,t==1), measurevar="SameRT", groupvars=c("roll_sum","monkey","Phase"))
Sumt1dif <- summarySE(subset(TrapSDif,t==1), measurevar="SameRT", groupvars=c("roll_sum","monkey","Phase"))

ggplot(Sumt1same, aes(x=roll_sum,y=SameRT)) + #What data to plot
  geom_point(aes(colour=monkey), size=3,alpha = 0.5)+
  geom_line(aes(colour=monkey))+
  geom_errorbar(aes(ymin=SameRT-se, ymax=SameRT+se), width=.1) +
    ylim(0,1) +
    geom_hline(aes(yintercept=0.333), colour="#990000", linetype="dashed") +
    ggtitle("Same Stimulus as 'Negative' Trap on t0") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("Proportion repetition same response on t1 as 'Neg' Trap") +
    xlab("# Rewards in 3 trials before 'Neg' Trap") +
    facet_grid(monkey ~ Phase)+ 
    theme_bw()

ggplot(Sumt1dif, aes(x=roll_sum,y=SameRT)) + #What data to plot
  geom_point(aes(colour=monkey), size=3,alpha = 0.5)+
  geom_line(aes(colour=monkey)) +
  geom_errorbar(aes(ymin=SameRT-se, ymax=SameRT+se), width=.1) +
    ylim(0,1) +
    geom_hline(aes(yintercept=0.333), colour="#990000", linetype="dashed") +
    ggtitle("Different Stimulus from 'Negative' Trap on t0") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("Proportion repetition same response on t1 as 'Neg' Trap") +
    xlab("# Rewards in 3 trials before 'Neg' Trap") +
    facet_grid(monkey ~ Phase)+ 
    theme_bw()
```


Performance of the 2 monkeys across the 3 stages of this task is fairly similar and strong. 
The monkeys look very much like the "good learner" humans in the PD manuscript, without the AI lesion
There is possibly a reduction in the trap reactivity across phases, so will study this below. 

So how does the monkeys' within-TS learning change across the stages?    


```{r warning = FALSE,echo = FALSE }

# Counter in each TS
Trials <- Trials %>%
  group_by(monkey,Phase,session, TS) %>% 
  dplyr::mutate(TScount = row_number(), TSperc = mean(COR))
#Summarise % correct
TSsum <- summarySE(Trials, measurevar="COR", groupvars=c("TScount","monkey","Phase"))
#Plot % correct through TS
ggplot(TSsum, aes(x=TScount, y=COR)) +
  geom_point(aes(colour=monkey), size=3,alpha = 0.5)+
  geom_line(aes(colour=monkey)) +
  geom_errorbar(aes(ymin=COR-se, ymax=COR+se), width=.1) +
  geom_hline(aes(yintercept=0.333), colour="#990000", linetype="dashed") +
  xlim(0,50) +
  facet_grid(monkey ~ Phase)+ 
  theme_bw()

#How well are the TSs solved? Is there a pattern over time?
TSs <- Trials %>%
  dplyr::group_by(monkey,Phase,session, TS) %>%
  slice(1)
  
  ggplot(TSs, aes(x=TSperc, fill = monkey)) +
  geom_histogram(position="dodge")+
  facet_grid(monkey ~ Phase)+ 
  theme_bw()

#Over time
TSs <- TSs %>%
    dplyr::group_by(monkey,Phase) %>%
    dplyr::mutate(TSf = as.factor(row_number()))

ggplot(TSs, aes(x=TSf, y=TSperc)) +
  geom_point(aes(colour=monkey), size=1)+
  facet_grid(monkey ~ Phase)+ 
  theme_bw()


#Look at only the good TSs
GoodTS <- subset(Trials, TSperc>0.75)
GoodTSsum <- summarySE(GoodTS, measurevar="COR", groupvars=c("TScount","monkey","Phase"))
ggplot(GoodTSsum, aes(x=TScount, y=COR)) +
  geom_point(aes(colour=monkey), size=3,alpha = 0.5)+
  geom_line(aes(colour=monkey)) +
  geom_errorbar(aes(ymin=COR-se, ymax=COR+se), width=.1) +
  geom_hline(aes(yintercept=0.333), colour="#990000", linetype="dashed") +
  xlim(0,50) +
  facet_grid(monkey ~ Phase)+ 
  theme_bw()

#Look at only the bad TSs
BadTS <- subset(Trials, TSperc<0.75)
BadTSsum <- summarySE(BadTS, measurevar="COR", groupvars=c("TScount","monkey","Phase"))
ggplot(BadTSsum, aes(x=TScount, y=COR)) +
  geom_point(aes(colour=monkey), size=3,alpha = 0.5)+
  geom_line(aes(colour=monkey)) +
  geom_errorbar(aes(ymin=COR-se, ymax=COR+se), width=.1) +
  geom_hline(aes(yintercept=0.333), colour="#990000", linetype="dashed") +
  xlim(0,50) +
  facet_grid(monkey ~ Phase)+ 
  theme_bw()

```

The comparison here is tricky as there is a criterion in ST1 but not afterwards. 
Maybe learning is initially faster in later STs? But overall there is a wider spread and more bad TSs in the later STs


Now we want to reproduce some of the figures from Faraut et al 2016, but for the current dataset

```{r warning = FALSE,echo = FALSE}

#Let's look at trap reactivity as %COR (Fig 4A Faraut)

#Need to consider both Trap types, so need a Trap count that does that

# Count trials since and to the Gtraps

Trials <- Trials %>%
  group_by(monkey, Phase) %>% 
  mutate (Traps = cumsum(GTrap+BTrap))
#max(Trials$Traps)
#sum(Trials$GTrap + Trials$BTrap)

Trials <- Trials %>%
  group_by(monkey, Phase, Traps) %>% 
  dplyr::mutate(TrapC = row_number()-1,TrapCback = row_number() - (n()+1))

# Subset down to trials just around Traps and set t around Traps
#rm ('Traps', 'TrapSE')
AllTraps <- subset(Trials, TrapC <=4 | TrapCback >= -4)
AllTraps <- transform(AllTraps, tt = ifelse(TrapC>4,TrapCback,TrapC))

TrapReac <- summarySE(AllTraps, measurevar="COR", groupvars=c("tt","monkey","Phase"))

ggplot(TrapReac, aes(x=tt,y=COR)) + #What data to plot
    geom_point(aes(color=monkey),size=3,alpha=0.5)+
    geom_line(aes(color=monkey)) +
    geom_errorbar(aes(ymin=COR-se, ymax=COR+se), width=.1) +
    ylim(0.3,0.7) +
    geom_hline(aes(yintercept=0.333), colour="#990000", linetype="dashed") +
    geom_rect(aes(xmin=-0.1, xmax=0.1, ymin=0.3, ymax=0.7), color="grey", alpha=0.01) +
    ggtitle("Trap Reactivity") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("% Correct") +
    xlab("Trials around Trap")+
    facet_grid(monkey ~ Phase)+ 
    theme_bw()

```

So in the early version the Trap reactivity is greater than it is more recently, but there are pre-trap differences between the monkeys so interpretation is hard. 
To be discussed....

So can we track a trap reactivity measure across the stages?


```{r warning = FALSE,echo = FALSE}

#Trials 0 and 1

TrapRe <- subset(AllTraps, tt==0 | tt==1)
TrapReS <- TrapRe %>%
  dplyr::group_by(monkey,Phase,session,tt) %>% #Group on tt to separate the outcomes
  dplyr::summarise(Cor=mean(COR)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(diff=Cor-lag(Cor,default=first(Cor))) # This calculates the trap reactivity
TrapReS <- subset(TrapReS, tt==1) #Fudge to only have the reactivity as 0 -> 1
TrapReS$diff <- TrapReS$diff * -1 # Trap reactivity as +ve value

ggplot(TrapReS, aes(x=session,y=diff)) + #What data to plot
    geom_point(aes(color=monkey),size=1,alpha=0.5)+
    geom_smooth(aes(color=monkey)) +
    ggtitle("Trap Reactivity Change over time") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("Trap reactivity") +
    xlab("Session")+
    facet_grid(monkey ~ Phase)+ 
    theme_bw()

```

The evidence for a diminishing trap reactivity over time is present but weak...

So what else can we add to this?
 